---
title: "Confidence Interval Methods"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Confidence Interval Methods}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

```{r setup}
library(TrialEmulation)
```

# Confidence Interval Methods

This vignette demonstrates the confidence interval methods available for per-protocol effect estimation and inference.

Confidence intervals are obtained through the `predict()` method for fitted `trial_sequence` objects. The confidence interval method is controlled through the `ci_type` argument.

Currently available methods are:

- `"sandwich"`
- `"Nonpara. bootstrap"`
- `"LEF outcome"`
- `"LEF both"`
- `"Jackknife Wald"`
- `"Jackknife MVN"`

The default method uses the sandwich variance estimator. Alternative methods based on bootstrap and jackknife resampling can be useful in settings with smaller sample sizes, sparse outcomes, or low treatment prevalence [1].

# Example Analysis

We begin by fitting a simple per-protocol target trial emulation.

## Create a Trial Sequence Object

```{r}
trial_pp <- trial_sequence(estimand = "PP")

trial_pp_dir <- file.path(tempdir(), "trial_pp")
dir.create(trial_pp_dir)

data("data_censored")

trial_pp <- trial_pp |>
  set_data(
    data = data_censored,
    id = "id",
    period = "period",
    treatment = "treatment",
    outcome = "outcome",
    eligible = "eligible"
  ) |>
  set_switch_weight_model(
    numerator = ~age,
    denominator = ~age + x1 + x3,
    model_fitter = stats_glm_logit(
      save_path = file.path(trial_pp_dir, "switch_models")
    )
  ) |>
  set_censor_weight_model(
    censor_event = "censored",
    numerator = ~x2,
    denominator = ~x2 + x1,
    pool_models = "none",
    model_fitter = stats_glm_logit(
      save_path = file.path(trial_pp_dir, "censor_models")
    )
  )|>
  calculate_weights() |>
  set_outcome_model() |>
    set_expansion_options(
    output = save_to_datatable(),
    chunk_size = 500
  ) |>
  expand_trials() |>
  load_expanded_data() |>
  fit_msm()
```

# Sandwich Confidence Intervals

The sandwich variance estimator is the default confidence interval method. It uses the robust variance-covariance matrix estimate of the marginal structural model parameters together with multivariate Normal sampling.

```{r}
pred_sandwich <- predict(
  trial_pp,
  newdata = outcome_data(trial_pp)[trial_period == 1, ],
  predict_times = 0:10,
  type = "survival",
  conf_int = TRUE,
  ci_type = "sandwich"
)

head(pred_sandwich$difference)
```

This is the fastest confidence interval method and is often suitable for larger datasets.

# Nonparametric Bootstrap

The nonparametric bootstrap repeatedly resamples patients and refits the complete analysis in each bootstrap sample.

```{r}
pred_boot <- predict(
  trial_pp,
  newdata = outcome_data(trial_pp)[trial_period == 1, ],
  predict_times = 0:10,
  type = "survival",
  conf_int = TRUE,
  ci_type = "Nonpara. bootstrap",
  samples = 100
)

head(pred_boot)
```

The `samples` argument specifies the number of bootstrap resamples.

# LEF Bootstrap Methods

The linearised estimating function (LEF) bootstrap methods provide computationally efficient alternatives to the ordinary bootstrap.

## LEF Outcome

The `"LEF outcome"` method uses a linear approximation for the marginal structural model parameters while refitting the weight models within each bootstrap sample.

```{r}
pred_lef_outcome <- predict(
  trial_pp,
  newdata = outcome_data(trial_pp)[trial_period == 1, ],
  predict_times = 0:10,
  type = "survival",
  conf_int = TRUE,
  ci_type = "LEF outcome",
  samples = 100
)

head(pred_lef_outcome)
```

## LEF Both

The `"LEF both"` method uses linear approximations for both the weight model parameters and the marginal structural model model parameters.

```{r}
pred_lef_both <- predict(
  trial_pp,
  newdata = outcome_data(trial_pp)[trial_period == 1, ],
  predict_times = 0:10,
  type = "survival",
  conf_int = TRUE,
  ci_type = "LEF both",
  samples = 100
)

head(pred_lef_both)
```

Because repeated model fitting is avoided, this method is often substantially faster than the ordinary nonparametric bootstrap.

# Jackknife Methods

The package also provides two jackknife-based confidence interval methods.

## Jackknife Wald

The Wald jackknife method estimates the standard error of the cumulative risk/survival difference using leave-one-patient-out resampling and constructs a Normal approximation confidence interval.

```{r}
pred_jack_wald <- predict(
  trial_pp,
  newdata = outcome_data(trial_pp)[trial_period == 1, ],
  predict_times = 0:10,
  type = "survival",
  ci_type = "Jackknife Wald"
)

head(pred_jack_wald)
```

## Jackknife MVN

The multivariate Normal jackknife method estimates a jackknife variance-covariance matrix for the outcome model parameters and uses multivariate Normal sampling, much like the sandwich variance-based confidence interval method.

```{r}
pred_jack_mvn <- predict(
  trial_pp,
  newdata = outcome_data(trial_pp)[trial_period == 1, ],
  predict_times = 0:10,
  type = "survival",
  conf_int = TRUE,
  ci_type = "Jackknife MVN"
)

head(pred_jack_mvn$difference)
```


The choice of confidence interval method depends on the size and characteristics of the dataset [1].

| Scenario | Recommended CI Method | Reason |
|----------|----------------------|---------|
| Small/medium sample size, low/medium event rate, low/medium treatment prevalence | LEF Bootstrap CI | Most reliable overall performance in sparse data settings. |
| Large sample size and medium/high event rate | Sandwich CI | Good performance with lowest computational cost. |
| Small sample size and low treatment prevalence | Jackknife Wald CI | Maintained close-to-nominal coverage. |
| Small sample size, medium/high event rate, and low/medium treatment prevalence | Jackknife MVN CI | Maintained close-to-nominal coverage. |

The LEF methods were developed to improve confidence interval performance in settings where sandwich-based intervals may exhibit undercoverage, while avoiding the computational burden of a full nonparametric bootstrap. The latter method can also be unstable in settings with small sample size and very low outcome event rates, where some bootstrap samples may result in non-invertible covariate matrices. In such cases, the lack of refitting in the LEF methods are beneficial. More discussion is given in the reference below [1].

# Additional Notes

- Resampling-based confidence intervals are currently available only for per-protocol estimands.
- The sandwich estimator remains the default confidence interval method.
- For bootstrap-based methods, the `samples` argument specifies the number of resamples. For `"sandwich"` and `"Jackknife MVN"` methods, the `samples` argument determines the sample size of the mutivariate Normal sampling.
- Larger values of `samples` improve Monte Carlo precision but increase computation time.
- Confidence intervals are calculated pointwise at each prediction time.

# Reference

[1] Limozin, J.M., Seaman, S.R. and Su, L., 2025. Inference procedures in sequential trial emulation with survival outcomes: Comparing confidence intervals based on the sandwich variance estimator, bootstrap and jackknife. Statistical Methods in Medical Research, 34(10), pp.2011-2033.

