---
title: "Extending-TrialEmulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending-TrialEmulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TrialEmulation)
```

Due to the extensive use of classes, TrialEmulation can be expanded by the user to fit their own specific needs.

This document gives a quick overview of the extensible classes, the current implementations and the requirements for
adding your own child classes.

## Model fitters

## Data Stores

In order to add a new data store, a child class must be defined that inherits from te_datastore.

You must also add at least a new constructor `save_to_xxx()` as well as new methods for `save_expanded_data()`
and `read_expanded_data()`. 

A new method for `sample_expanded_data()` is optional (e.g. in case sampling is not required or the implemented method
for te_datastore is sufficient, see below under *sample_expanded_data*), but it might be necessary for large datasets.

### Classes and Slots

- **te_datastore**: Parent class, placed as a place holder in trial_sequence objects before setting expansion options, 
will be replaced with the corresponding child class when expansion options are set.
  - @N: Number of observations

Currently the following Data Store child classes are available for saving expanded data:

- **te_datastore_csv**: Expanded data is saved as csv files, one file per trial period. When reading the data, only the
files corresponding to the selected trial periods are read.
  - @path: Path to temp folder containing the csv files
  - @files: Paths to all available files
  - @template: empty data.frame, used as a template when reading the data
  - @N: inherited from te_datastore
- **te_datastore_datatable**: Expanded data is saved as a data.table in memory, only viable for smaller datasets.
  - @data: data.table containing expanded data
  - @N: inherited from te_datastore
- **te_datastore_duckdb**: Expanded data is saved as a duckdb file containing all trial periods. Reading, subsetting and
sampling can be done efficiently with an SQL query (currently constructed with a translator helper function).
  - @path: Path to temp folder containing the duckdb file
  - @table:
  - @con: A duckdb_connection object, used to connect to the file in a query
  - @N: inherited from te_datastore




### Constructor

A constructor function is required to be used in `set_expansion_options()` to replace the te_datastore object in
`trial_sequence@expansion@datastore` with an object of the corresponding child class. Saving of the data happens when
calling `expand_trials()` which internally calls the corresponding `save_expanded_data()` method.

See the following currently available constructor functions for further insights: 
`save_to_csv()`, `save_to_datatable()`, `save_to_duckdb()`

### Methods

#### show

*Note:* Since the child classes differ quite significantly from each other, every child class has its own show method. 
There is no show method for the te_datastore parent class.

#### save_expanded_data

This method defines how the expanded data gets saved. Method is chosen based on the te_datastore child class.
It gets called internally by `expand_trials()`.

- **Arguments**
  - object: a te_datastore child class object
  - data: data to be saved
- **Returns**: a modified te_datastore child class object

#### read_expanded_data

This method is used for reading the expanded data into memory. The data can be subset by period or any other subset
condition. It gets called internally by `load_expanded_data()` if p_control isn't specified and by
`sample_expanded_data()` if no specific sampling method exists for a te_datastore child class.

- **Arguments**
  - object: a te_datastore child class object
  - period: integerish vector to select trial periods, if missing defaults to `NULL` and selects all available trial
  periods
  - subset_condition: subset condition as a string, if missing defaults to `NULL` and skips subsetting
- **Returns**: a data.table object

#### sample_expanded_data

This method is used for reading and sampling the expanded data. The data can be subset by period or any other subset
condition plus it can be sampled using the p_control argument. It gets called internally by `load_expanded_data()` if
p_control is specified. 

If no method for the child class exists, the method of the parent class will be used instead which will read and subset
the data using `read_expanded_data()`. Then the sampling happens in bulk, which might cause problems for large datasets.
Because of memory limitations it might be necessary to implement a more efficient method for a new child class.

- **Arguments**
  - object: a te_datastore child class object
  - p_control: numeric value between 0 and 1, probability to sample a control value
  - period: integerish vector to select trial periods, if missing defaults to `NULL` and selects all available trial
  periods
  - subset_condition: subset condition as a string, if missing defaults to `NULL` and skips subsetting
  - seed: a seed to be used for sampling, if missing sampling is randomised
- **Returns**: a data.table object
