---
title: "Getting-Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting-Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bigmemory)
options(bigmemory.default.shared = FALSE)
```

This package implements methods to analyse a sequence of target trials.

The steps to do this are:

 1. create the data set with all the required variables
 2. fit models for censoring and calculate weights
 3. expand the data set, with records for each eligible patient for each trial period
 4. fit models for outcomes

Some of the functions are inspired by the [INITIATORS SAS macro](https://causalab.sph.harvard.edu/software/).

```{r setup}
library(RandomisedTrialsEmulation)
```


There is an all-in-one function `initators()` which does all of the steps with one function call.

Currently, datasets are handled with csv files[^csv]. 

To get started a longitudinal dataset must be created and saved as a csv.
This should have:

 * time period
 * patient identifier
 * treatment indicator
 * outcome indicator
 * eligibility indicator for trial start in period
 * other covariates relating to treatment, outcome, or censoring to be used in the models for weights or the outcome
 

[^csv]: Saving the data to disk as csv allows us to use various techniques when the expanded data set becomes very large and no longer fit easily in memory.


```{r}
# Prepare the example data
data("trial_example")

# for the purposes of the vignette, we use a temporary directory, however it may be useful to use a permanent location in order to inspect the outputs later
working_dir <- file.path(tempdir(),"trial_emu") 
dir.create(working_dir)
data_path <- file.path(working_dir, "trial_example.csv")
write.csv(trial_example, file=data_path)
```

Call the `initiators()` function to run the complete analysis:
```{r}
result <-
  initiators(data_path = data_path,
             id="id",
             period = "period",
             eligible = "eligible",
             treatment = "treatment",
             outcome = "outcome",
             model_var = "assigned_treatment",
             outcomeCov_var = c("catvarA", "catvarB", "catvarC","nvarA","nvarB","nvarC"),
             outcomeCov =  c("catvarA", "catvarB", "catvarC","nvarA","nvarB","nvarC"),
             outcomeClass = c("catvarA", "catvarB", "catvarC"),
             numCores = 1,
             data_dir = working_dir,
             use_censor = 0,
             use_weight = 0)
```

`initiators()` creates a few additional data files, depending on the selected options
```{r}
list.files(working_dir)
```
 
 * `sw_data` is created in the first step and contains all the necessary variables for the next steps
 * `switch_data` is the extended data
 

Instead of using `initiators()` it is possible to run each step individually with `data_preparation()`, `data_modelling()` etc.

```{r, eval = FALSE}
model_result <- data_modelling(
  outcomeCov_var = c("catvarA", "catvarB", "catvarC","nvarA","nvarB","nvarC"),
  model_var = "assigned_treatment",
  include_followup_time_case = "linear",
  include_expansion_time_case = "linear",
  absolutePath = file.path(working_dir, "switch_data.csv"),
  use_sample_weights = FALSE,
  quiet = TRUE
)
```



```{r, echo=FALSE}
# clean up
unlink(working_dir)
```



