---
title: "Getting-Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting-Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package implements methods to analyse a sequence of target trials.

The steps to do this are:

 1. create the data set with all the required variables
 2. fit models for censoring and calculate weights
 3. expand the data set, with records for each eligible patient for each trial period
 4. fit models for outcomes

Some of the functions are inspired by the [INITIATORS SAS macro](https://causalab.sph.harvard.edu/software/).

```{r setup}
library(TrialEmulation)
```


There is an all-in-one function `initators()` which does all of the steps with one function call.

To get started a longitudinal dataset must be created containing:

 * time period
 * patient identifier
 * treatment indicator
 * outcome indicator
 * eligibility indicator for trial start in period
 * other covariates relating to treatment, outcome, or censoring to be used in the models for weights or the outcome
 

```{r data}
# Prepare the example data
data("trial_example")

trial_example$catvarA <- as.factor(trial_example$catvarA)
trial_example$catvarB <- as.factor(trial_example$catvarB)
trial_example$catvarC <- as.factor(trial_example$catvarC)

head(trial_example)
```

Call the `initiators()` function to run the complete analysis:
```{r initiators}
result <- initiators(
  data = trial_example,
  id = "id",
  period = "period",
  eligible = "eligible",
  treatment = "treatment",
  outcome = "outcome",
  model_var = "assigned_treatment",
  outcome_cov = c("catvarA", "catvarB", "catvarC", "nvarA", "nvarB", "nvarC"),
  use_censor = 0,
  use_weight = 0
)
```

By default, this returns the final `glm` model object and the results using the sandwich estimator.
```{r init_summary}
summary(result$model)
```

```{r init_robust_summary}
result$robust$summary
```

Also the sandwich variance-covariance matrix.
```{r init_robust_matrix}
result$robust$matrix
```


# Flexible Analysis

To gain complete control over the analysis and to inspect the intermediate objects, it can
be useful to run the the data preparation and modelling steps separately.

This also allows for processing of very large data sets, by doing the data preparation in chunks and then
sampling from the expanded trial data.

```{r temp_dir}
# for the purposes of the vignette, we use a temporary directory, however it may be useful to use a permanent location
# in order to inspect the outputs later
working_dir <- file.path(tempdir(TRUE), "trial_emu")
if (!dir.exists(working_dir)) dir.create(working_dir)
```

```{r data_preparation}
prep_data <- data_preparation(
  data = trial_example,
  id = "id",
  period = "period",
  eligible = "eligible",
  treatment = "treatment",
  outcome = "outcome",
  model_var = "assigned_treatment",
  outcome_cov = ~ catvarA + catvarB + catvarC + nvarA + nvarB + nvarC,
  data_dir = working_dir,
  save_weight_models = TRUE,
  use_censor = 1,
  use_weight = 1,
  chunk_size = 500,
  separate_files = TRUE,
  switch_n_cov = ~ nvarA + nvarB
)
```
We have some summaries of the weighting models in the results from `data_preparation()`
```{r weight_summaries}
str(prep_data$switch_models, max.level = 2)
prep_data$switch_models$switch_n0
prep_data$switch_models$switch_n0_statistics
```

We also have the full model objects saved in `data_dir = working_dir`, so you can inspect the data used in those models and further investigate how well those models fit.
```{r weight_files}
list.files(working_dir, "*.rds")

switch_n0 <- readRDS(file.path(working_dir, "weight_model_switch_n0.rds"))
summary(switch_n0)
```

We also see the expanded trial files:
```{r trial_files}
list.files(working_dir, pattern = "trial_")
```

Each of these csv files contains the data for the trial starting at period `_i`.
To create a manageable dataset for the final analysis we can sample from these trials.

Here we sample 10% of the patients without an event at each follow-up time in each trial. All 
observations with events are included.
```{r sample}
sampled_data <- case_control_sampling_trials(prep_data, p_control = 0.1, sample_all_times = TRUE)
str(sampled_data)
```

Before proceeding with the modelling, it is possible to manipulate and derive new variables and adjust factor levels
in the `data.frame`. You can also specify transformations in a formula to `outcome_cov`, 
`include_followup_time_case` or `include_expansion_time_case`, such as `~ ns(for_period)` or 
`~ I(nvarA^2) + nvarC > 50 + catvarA:nvarC`.

It is also possible to specify formulas in `data_modelling` to include, for example, splines with
`ns()`.

Now we can fit the model with the `data_modelling()` function. Since we have sampled from the data, 
we should make use of the `use_sample_weights` option to get correct survival estimates.

```{r modelling}
model_result <- data_modelling(
  data = sampled_data,
  outcome_cov = c("catvarA", "catvarB", "catvarC", "nvarA", "nvarB", "nvarC"),
  model_var = "assigned_treatment",
  use_weight = 1,
  use_censor = 0,
  glm_function = "glm",
  use_sample_weights = TRUE
)
```

The result is the same type as the previous result from the simple `initiators` function, 
containing the `glm` object and the sandwich results.

```{r modelling_result}
str(model_result, max.level = 1)
```

```{r glm_summary}
summary(model_result$model)
```

We can also use this object to predict cumulative incidence curves and use these for treatment comparisons.
Here we use the patients who were in the first trial as the target population. To get the data in the right format,
we can use the `data_template` returned by `data_preparation()`.

```{r predict}
new_data <- data.table::fread(file.path(working_dir, "trial_1.csv"))
new_data <- rbind(data.table::as.data.table(prep_data$data_template), new_data)
model_preds <- predict(model_result, predict_times = c(0:40), newdata = new_data, type = "cum_inc")
```

The `predict` function returns a list of 3 data frames: predictions for assigned treatment 0, for assigned treatment 1
and for the difference. It possible to change the type of predicted values, either cumulative incidence or survival.

```{r}
plot(
  model_preds$difference$followup_time,
  model_preds$difference$cum_inc_diff,
  ty = "l", ylab = "Cumulative Incidence Difference",
  xlab = "Follow-up Time",
  ylim = c(-0.15, 0.02)
)
lines(model_preds$difference$followup_time, model_preds$difference$`2.5%`, lty = 2)
lines(model_preds$difference$followup_time, model_preds$difference$`97.5%`, lty = 2)
```


```{r cleanup, echo=FALSE}
# clean up
unlink(working_dir, recursive = TRUE)
```
