---
title: "New Interface"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{New Interface}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TrialEmulation)
```

To improve usability, we have implemented a new user interface. This allows a more structured specification of the 
target trial emulation process.

It also gives flexibility to add new methods and tools for parts of the analysis. For example, we now allow different
ways of storing the expanded data: as CSV files and in a DuckDB database. We also allow different weight fitting model
procedures: using `stats::glm` or `parglm::parglm`. New components can quickly and easily be specified for use with
this package.


## User Interface

A sequence of target trials analysis starts by specifying which estimand will be used:

```{r}
trial_pp <- trial_sequence(estimand = "PP") # Per-protocol
trial_itt <- trial_sequence(estimand = "ITT") # Intention-to-treat
```

Additionally it is useful to create a directory to save files for later inspection.
```{r}
trial_pp_dir <- file.path(tempdir(), "trial_pp")
dir.create(trial_pp_dir)
trial_itt_dir <- file.path(tempdir(), "trial_itt")
dir.create(trial_itt_dir)
```


### Data

Next the user must specify the data that will be used. Here we need to specify which columns contain which values
and how they should be used.
```{r}
data("data_censored")
trial_pp <- trial_pp |>
  set_data(
    data = data_censored,
    id = "id",
    period = "period",
    treatment = "treatment",
    outcome = "outcome",
    eligible = "eligible",
    expand_variables = c("x1", "x2", "x3", "x4", "age", "age_s")
  )

# Function style without pipes
trial_itt <- set_data(
  trial_itt,
  data = data_censored,
  id = "id",
  period = "period",
  treatment = "treatment",
  outcome = "outcome",
  eligible = "eligible",
  expand_variables = c("x1", "x2", "x3", "x4", "age", "age_s")
)
```

We can inspect our object by printing:
```{r}
trial_itt
```


### Weight Models

To adjust for the effects of informative censoring, inverse probability of censoring weights (IPCW) can be applied. To
estimate these weights, we construct survival models. Two sets of models are fit for the two censoring mechanisms which
may apply: censoring due to deviation from assigned treatment, and other informative censoring.

#### Censoring due to treatment switching

We specify model formulas to be used for calculating the probability of receiving treatment in the current period.
Separate models are fitted for patients who had `treatment = 1` and those who had `treatment = 0` in the previous 
period. Stabilized weights are used by fitting numerator and denominator models.

There are optional arguments to specify columns which can include/exclude observations from the treatment models. These
are used in case it is not possible for a patient to deviate from a certain treatment assignment in that period.

```{r}
trial_pp <- trial_pp |>
  set_switch_weight_model(
    numerator = ~ age + x1 + x3,
    denominator = ~age,
    model_fitter = stats_glm_logit(save_path = file.path(trial_pp_dir, "switch_models"))
  )
trial_pp
```

This type of censoring is not used with an ITT estimand, so we cannot use `set_switch_weight_model()` with `trial_ITT` objects.

#### Other informative censoring

In case there is other informative censoring occurring in the data, we can create similar models to estimate the IPCW.
These can be used with all types of estimand. 
Compared to `set_switch_weight_model` there are additional required arguments:

 - `censor_event` which specifies the column containing the censoring indicator
 - `pool_models` which species that models may be fit separately (as in `set_switch_weight_model`) or pooled across the
treatments in the previous period. The choices are `"none"`, `"both"`, or `"numerator"` only. The default and allowed
choices depends on the estimand.


```{r}
trial_pp <- trial_pp |>
  set_censor_weight_model(
    censor_event = "censored",
    numerator = ~ x1 + x2 + x3,
    denominator = ~x2,
    pool_models = "none",
    model_fitter = stats_glm_logit(save_path = file.path(trial_pp_dir, "switch_models"))
  )
trial_pp
```

```{r}
trial_itt <- set_censor_weight_model(
  trial_itt,
  censor_event = "censored",
  numerator = ~ x1 + x2 + x3,
  denominator = ~x2,
  pool_models = "numerator",
  model_fitter = stats_glm_logit(save_path = file.path(trial_pp_dir, "switch_models"))
)
trial_itt
```

#### Calculate Weights

Next we need to fit the individual models and combine them into weights. This is done with `calculate_weights()`.

```{r}
trial_pp <- trial_pp |> calculate_weights()
trial_pp


trial_itt <- calculate_weights(trial_itt)
```

The full model objects are saved to disk in the directories we created above. The summaries are stored in the trial
sequence object and can be printed:

```{r}
show_weight_models(trial_itt)
```

### Specify Outcome Model

 Not fully implemented. Needs to be preliminarily set at this stage to make sure all required variables are included
 in the expanded data set
```{r, eval = FALSE}
trial_pp <- set_outcome_model(trial_pp)
trial_itt <- set_outcome_model(trial_itt, adjustment_terms = ~ x1 + x2)
```
 

### Expand Trials

Now we are ready to create the data set with all of the sequence of target trials. First we specify some options for the
expansion and then expand.

#### Set Expansion Options

There are two options to set

 - output: specifies how and where the expanded data will be saved. As it can be very large, we may need to save it to
 disk with CSV files or DuckDB, using a `save_to_*` function.
 - chunk_size: if the expanded data is too large to fit in memory, we need to process it in chunks by specifying how
 many patients are processed at one time.

```{r}
trial_pp <- trial_pp |>
  set_expansion_options(output = save_to_csv(file.path(trial_pp_dir, "trial_csvs")), chunk_size = 500)

trial_itt <- set_expansion_options(
  trial_itt,
  output = save_to_csv(file.path(trial_itt_dir, "trial_csvs")), chunk_size = 500
)

trial_itt <- set_expansion_options(
  trial_itt,
  output = save_to_data.table(), chunk_size = 500
)
```


#### Create Sequence of Trials Data

```{r}
trial_pp <- expand_trials(trial_pp)
trial_itt <- expand_trials(trial_itt)
```


Show method not implemented yet
```{r}
trial_pp@expansion
trial_itt@expansion
```


### Sample from Expanded Data

Sample method not implemented yet

### Fit Marginal Structural Model

Outcome model fitting not implemented yet

### Inference

Model summaries not implemented

## Class Structure

The main class in this new interface is `trial_sequence` which is the parent class of the estimand specific classes
`trial_sequence_ITT`, `trial_sequence_PP` and `trial_sequence_AT`. These child classes allow for restrictions and 
differences in processing based on the different estimands. Where necessary we have S4 methods defined for the estimand
specific classes and otherwise common processing is defined in methods for the `trial_sequence` class.

The `trial_sequence` class contains slots needed to define the sequence of target trials analysis

 - estimand = `character`
 - data = `te_data`
 - censor_weights = `te_weights_spec`
 - expansion = `te_expansion`
 - outcome_model = `te_outcome_model`


### `te_data`

Contains the input data. It under goes some manipulation in `set_data()` to standardise the column names.

### `te_weights_spec`

Contains specification for the IPCW models, including formulas, model fitter, pooling option, and (once fitted) the 
model summaries. 

#### te_weights_fitter

The model fitter is specified as an object of class `te_weights_fitter`. This class is used to save parameters which are
used in the method `fit_weights_model(object, data, formula, label)` where object is the `te_weights_fitter`.
It fits a model and returns a `te_weights_fitted` object containing the label, model summaries and the fitted values 
from the model.

#### te_weights_fitted

The `te_weights_fitted` objects are saved in a list in `te_weights_spec` in a slot `fitted`.

### `te_outcome_model`

To be finalised.

### `te_expansion`

The `te_expansion` class defines how the sequence of trials should be expanded. It contains slots for chunk_size,
how censoring is applied, first and last periods, and an object of class `te_datastore` which defines how the data is
saved.

#### `te_datastore`

This class contains slots needed for saving the expanded data, such as file path. Child classes are defined for each
storage type, such as `te_datastore_csv`, `te_datastore_duckdb`. Each of these classes has a user-friendly 
`save_to_*()` method to setup the storage.

To define a new storage type, a method `save_expanded_data(object, data)` needs to be defined. This takes the data store
object and a `data.frame` to be saved. This method may be called repeatedly if only a chunk of patients are being
processed and so should appropriately append the data. The method returns an updated `te_datastore_*` object which is
saved in the `trial_sequence@te_expansion` slot.
