---
title: "outcome_data_model"
output: html_document
date: "2024-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## User interface

After expansion, we need to fit the outcome model. Before that we need to prepare the outcome data set.
This might involve sampling, or some other filtering or variable derivation.

```{r}
trial_itt <- trial_sequence(...) |> ... |> set_outcome_model(...) |> expand_trials()
```

### simple case

with small data it should just work

```{r}
fit_msm(trial_itt)
```

This would load the full expanded data and fit the model

### with sampling

This should be done simply
```{r}
trial_itt <- sample_controls(trial_itt, p_control = 0.3, seed = 444)
fit_msm(trial_itt)
```

### some complex filtering

```{r}
model_data(trial_itt) <- model_data(trial_itt) |> group_by() |> mutate()
fit_msm(trial_itt)
```


## Internal Representation

Currently the input data has a slot `@data` and a class `te_data` which has some validation, eg on
columns.

We could have a second slot eg `@model_data`. We would need a different or modified `te_data` class, at
least the validation. Alternatively we could directly have the data.table in the slot and validation could
occur in the getter/setter methods.


### Option: data.table

Remove `te_data` class and do all validation in setter methods.

- lose class validation
- lose N and Nobs slots
- lose `te_data_unset` class
+ don't need `@data@data`

### Option: te_data classes

Modify class

- don't validate in class, but have a required columns slot
- can store vector of periods
- lose `te_data_unset` class

### Option: te_data_model, te_data_input

More subclasses which inherit from `te_data`

- natural in OO
- can have all validation built in
- have slots with N, nobs, period etc

## Internal flow

If we have the two data slots, `@data`, `@model_data`

- `sample_controls()` would save the data.table to this slot
- could have a getter/setter `model_data(object)`/`model_data(object)<-` to access `@model_data@data`
- and maybe the save for the input data, `input_data(object)`/`input_data(object)<-` to access `@data@data`
- If `trial_msm()` is called, but `@model_data` is empty, it will be read from the `te_datastore`.
   - do we need to check that expansion has been done? --> further consider different class post expansion
   - If `is(@model_data, "te_data_unset")` give error: run data expansion
   - If `is(@outcome_model, "te_outcome_unset")` give error: run set outcome model
